Automating Verification of MPC Security

### 2023-06-07 popl24/minicat.tex
checkpoint

\emph{Memories} are fundamental to the semantics of $\fedcat$ and
provide the random tape and secret inputs to protocols, and record
view assignments. Memories $\store$ are finite (partial) mapping from
variables to values. The \emph{domain} of a memory is written
$\dom(\store)$ and is the finite set of variables on which the memory
is defined. We write $\store\{ x \mapsto v\}$ for
$x\not\in\dom(\store)$ to denote the memory $\store'$ such that
$\store'(x) = v$ and otherwise $\store'(y) = \store(y)$ for all $y \in
\dom(\store)$.

@@

Given a variable-free boolean expression $\be$, we write $\cod{\be}$
to denote the standard interpretation of $\be$.  With the introduction
of variables to expressions, we have two concerns. First, we need to
interpret variables with respect to a specific memory, and second,
we need to ensure that variables are used ``legally''. That is, since
expressions define local computation, all variables used in an
expression must belong to the same client.  Thus, we denote
interpretation of expressions $\be$ possibly containing variables as
$\lcod{\store,\be}{\cid}$, where $\store$ associates variables with
values and all variables must be owned by client $\cid$. This is
defined in Figure \ref{fig-minifed-interp}.

\begin{fpfig}[t]{$\minifed$ expression interpretation.}{fig-minifed-interp}
\begin{eqnarray*}
\lcod{\store, \etrue}{\cid} &=& \etrue\\
\lcod{\store, \efalse}{\cid} &=& \efalse\\
\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
             \begin{cases}
                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
                \lcod{\store,\be_3}{\cid} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
             \end{cases}%\\
%\lcod{\store, \OT{\be_1}{\be_2}{\be_3}}{\cid} &=&
%             \begin{cases}
%                \lcod{\store,\be_2}{\cid'} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
%                \lcod{\store,\be_3}{\cid'} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
%             \end{cases}
\end{eqnarray*}
\end{fpfig}

Evaluation of configurations is then defined via a small-step reduction relation $\redx$.
This is defined in the obvious manner for view assignments other than through
$\ttt{OT}$- however note that reduction requires that views are never reassigned. 
\begin{mathpar}
  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)
\end{mathpar}
For view assignments through $\ttt{OT}$, we define a different reduction rule that
captures the appropriate semantics- that is, the selection bit is \emph{not} communicated
to the sender (through a view assignment), and only the selected bit is sent to the receiver:
\begin{mathpar}
  \inferrule
  {b = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{b}, \prog)}
\end{mathpar}
We define $\redxs$ as the reflexive, transitive closure of $\redx$.

>>>>

\emph{Memories} are fundamental to the semantics of $\fedcat$ and
provide the random tape and secret inputs to protocols, and record
view assignments. Memories $\store$ are finite (partial) mapping from
variables to binary values $\beta \in \{0,1\}$. The \emph{domain} of a
memory is written $\dom(\store)$ and is the finite set of variables on
which the memory is defined. We write $\store\{ x \mapsto \beta\}$ for
$x\not\in\dom(\store)$ to denote the memory $\store'$ such that
$\store'(x) = \beta$ and otherwise $\store'(y) = \store(y)$ for all $y \in
\dom(\store)$.

@@

Given a variable-free boolean expression $\be$, we write $\cod{\be}$
to denote the standard interpretation of $\be$ in the binary field.
With the introduction of variables to expressions, we have two
concerns. First, we need to interpret variables with respect to a
specific memory, and second, we need to ensure that variables are used
``legally''. That is, since expressions define local computation, all
variables used in an expression must belong to the same client.  Thus,
we denote interpretation of expressions $\be$ possibly containing
variables as $\lcod{\store,\be}{\cid}$, where $\store$ associates
variables with values and all variables must be owned by client
$\cid$. This is defined in Figure \ref{fig-minifed-interp}.

\begin{fpfig}[t]{$\minifed$ expression interpretation.}{fig-minifed-interp}
\begin{eqnarray*}
\lcod{\store, \etrue}{\cid} &=& 1\\
\lcod{\store, \efalse}{\cid} &=& 0\\
\lcod{\store, \flip{\cid}{w}}{\cid} &=& \store(\flip{\cid}{w})\\
\lcod{\store, \secret{\cid}{w}}{\cid} &=& \store(\secret{\cid}{w})\\
\lcod{\store, \view{\cid}{w}}{\cid} &=& \store(\view{\cid}{w})\\
\lcod{\store, \oracle{w}}{\cid} &=& \store(\oracle{w})\\
\lcod{\store, \enot\ \be}{\cid} &=& \cod{\enot\ \lcod{\store,\be}{\cid}}\\
\lcod{\store, \be_1\ \mathit{binop}\ \be_2}{\cid} &=&
    \cod{\lcod{\store,\be_1}{\cid}\ \mathit{binop}\ \lcod{\store,\be_2}{\cid}}\\
\lcod{\store, \select{\be_1}{\be_2}{\be_3}}{\cid} &=&
             \begin{cases}
                \lcod{\store,\be_2}{\cid} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
                \lcod{\store,\be_3}{\cid} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
             \end{cases}%\\
%\lcod{\store, \OT{\be_1}{\be_2}{\be_3}}{\cid} &=&
%             \begin{cases}
%                \lcod{\store,\be_2}{\cid'} & \text{if\ } \lcod{\store,\be_1}{\cid}\\
%                \lcod{\store,\be_3}{\cid'} & \text{if\ } \neg\lcod{\store,\be_1}{\cid}
%             \end{cases}
\end{eqnarray*}
\end{fpfig}

Evaluation of configurations is then defined via a small-step reduction relation $\redx$.
This is defined in the obvious manner for view assignments other than through
$\ttt{OT}$- however note that reduction requires that views are never reassigned. 
\begin{mathpar}
  (\store, \eassign{\view{\cid}{w}}{\be};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\lcod{\store,\be}{\cid'}}, \prog)
\end{mathpar}
For view assignments through $\ttt{OT}$, we define a different reduction rule that
captures the appropriate semantics- that is, the selection bit is \emph{not} communicated
to the sender (through a view assignment), and only the selected bit is sent to the receiver:
\begin{mathpar}
  \inferrule
  {\beta = \text{if\ } \lcod{\store,\be_1}{\cid}  \text{\ then\ } \lcod{\store,\be_2}{\cid'} \text{\ else\ } \lcod{\store,\be_3}{\cid'}}
      {(\store, \eassign{\view{\cid}{w}}{\OT{\be_1}{\be_2}{\be_3}};\prog) \redx (\extend{\store}{\view{\cid}{w}}{\beta}, \prog)}
\end{mathpar}
We define $\redxs$ as the reflexive, transitive closure of $\redx$.

### 2023-06-07 popl24/pmf.tex
checkpoint

\begin{definition}
  A \emph{memory probability distribution} $\pdf{X}$ is a function mapping
  memories in $\mems(X)$ to values in $[0..1]$, such that:
  $$
  \sum_{\store \in \mems(X)} \pdf{X}(\store) \  = \ 1
  $$
  We call $X$ the \emph{domain} of $\pdf{X}$.
\end{definition}

\begin{definition}
  Given $X$ the \emph{marginal distribution} of $Y \subseteq X$
  in a distribution $\pdf{X}$, denoted $\margd{\pdf{X}}{Y}$,
  is a distribution with domain $Y$ where for all
  $\store \in \mems(Y)$:
  $$
  (\margd{\pdf{X}}{Y})(\store) =
  \sum_{\store' \in \mems(X-Y)} \pdf{X}(\store \cup \store')
  $$
\end{definition}

\begin{definition}
  Given $\pdf{X}$, its \emph{conditional distribution given
  $\store$} with $\dom(\store) = Y$ and $Y \subseteq X$, denoted
  $\condd{\pdf{X}}{\store}$, is a distribution with domain $X - Y$ where for all
  $\store' \in \mems(X - Y)$:
  $$
  (\condd{\pdf{X}}{\store})(\store') =
      \pdf{X}(\store \cup \store') / ((\margd{\pdf{X}}{Y})(\store))
  $$
\end{definition}

Now we can define the probability distribution of a program $\prog$,
that we denote $\progd(\prog)$. Recall that $\fedcat$ is
deterministic, so a run is determined by the input secrets together
with random flip assignments. And since we constrain programs to not
overwrite memories, we can be sure that final memories resulting from
computation contain both a complete record of all initial secrets and
flips as well as views of communicated information. Thus a final
memory represents a run of the program, and since we consider and set
of \emph{input} secrets and flips to be equally likely, we consider
each memory in the set of final memories of a program to be equally
likely, which establishes the basic probability distribution for the
program. And since we don't care about flip values which are only
obversable insofar as they affect views, the probability distribution
we are really interested in is the marginal distribution of secrets
and views in this basic program distribution.
\begin{definition}
  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pdf{S \cup V \cup F}(\store) =
  \begin{cases}
    1 / |\stores| & \text{if}\ \store \in \stores\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pdf{S \cup V \cup F}}{S\cup V}
  $$
\end{definition}

>>>>

In order to precisely characterize distributions of secrets and views
in programs we define probability mass functions in a standard manner,
following the formulation in \ref{XXX}. In this formulation, memories
denote conjunctions of variable assignments. Intuitively, given a
program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
\mems(\iov(\prog))$, when considering the execution of $\prog$
initialized with secrets $\store_S$, we are fundamentally concerned
with the joint probability of the assignments in $\store_V$
conditioned by the assignments $\store_S$. This is because
probabilistic dependence between honest secrets and corrupt views
\emph{is} information leakage.

\subsection{Basic Definitions} We begin by defining memory probability distributions as follows.
\begin{definition}
  A \emph{memory probability distribution} $\pdf{X}$ is a function
  mapping memories in $\mems(X)$ to values in $[0..1]$, such that:
  $$
  \sum_{\store \in \mems(X)} \pdf{X}(\store) \  = \ 1
  $$
  We call $X$ the \emph{domain} of $\pdf{X}$.
\end{definition}
To recover succinct and familiar notation, we may omit the domain of a
distribution when it is clear from an application context-
i.e., we allow the following sugaring:
$$
\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $X$ the \emph{marginal distribution} of $Y \subseteq X$
  in a distribution $\pdf{X}$, denoted $\margd{\pdf{X}}{Y}$,
  is a distribution with domain $Y$ where for all
  $\store \in \mems(Y)$:
  $$
  (\margd{\pdf{X}}{Y})(\store) =
  \sum_{\store' \in \mems(X-Y)} \pdf{X}(\store \cup \store')
  $$
\end{definition}

\begin{definition}
  Given $\pdf{X}$, its \emph{conditional distribution given
  $\store$} with $\dom(\store) = Y$ and $Y \subseteq X$, denoted
  $\condd{\pdf{X}}{\store}$, is a distribution with domain $X - Y$ where for all
  $\store' \in \mems(X - Y)$:
  $$
  (\condd{\pdf{X}}{\store})(\store') =
      \pdf{X}(\store \cup \store') / ((\margd{\pdf{X}}{Y})(\store))
  $$
\end{definition}
To recover familiar notation we allow the following syntactic
sugarings:
\begin{eqnarray*}
\pdf{X}(\store|\store') &\defeq& (\condd{\pdf{X}}{\store})(\store')\\
\pdf{X}(\store) &\defeq& (\margd{\pdf{X}}{\dom(\store)})(\store) \qquad \dom(\store) \subset X
\end{eqnarray*}

\subsection{Program Distributions}
Now we can define the probability distribution of a program $\prog$,
that we denote $\progd(\prog)$. Since $\fedcat$ is deterministic the
results (view assignments) of any run are determined by the input
secrets together with the random tape. And since we constrain programs
to not overwrite views we are assured that \emph{final} memories
contain both a complete record of all initial secrets as well as views
resulting from communicated information.

We consider random tapes to be selected from a uniform distribution.
In our threat model we will also assume that any possible initial
secrets are equally likely when formulating the program distribution.
In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
V$ we will consider all $\store
\in \mems(\iov(\prog) \cup \flips(\prog))$ such that
$
\config{\store_{S \cup \flips(\prog)}}{\prog} \redxs \config{\store_}{\varnothing}
$
to be equally probable, establishing the basic distribution of the
program. From this we can immediately derive the marginal distribution
of $S \cup V$ to reason about dependencies between secrets and views. 
\begin{definition}
  \label{def-progd}
  Given progam $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  Define also $\pdf{S \cup V \cup F}$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pdf{S \cup V \cup F}(\store) =
  \begin{cases}
    1 / |\stores| & \text{if}\ \store \in \stores\\
    0 & \text{otherwise}
  \end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution:
  $$
  \progd(\prog) =  \margd{\pdf{S \cup V \cup F}}{S\cup V}
  $$
\end{definition}
